---
title: Project title
author: by Team name
output: github_document
---

## Summary

Write-up of your project and findings go here. 
Think of this as the text of your presentation. 
The length should be roughly 5 minutes when read out loud.
Although pacing varies, a 5-minute speech is roughly 750 words. 
To use the word count addin, select the text you want to count the words of (probably this is the 
Summary section of this document, go to Addins, and select the `Word count` addin).
This addin counts words using two different algorithms, but the results should be similar and as long as you're in the ballpark of 750 words, you're good! 
The addin will ignore code chunks and only count the words in prose.

You can also load your data here and present any analysis results / plots, but I strongly urge you to keep that to a minimum (maybe only the most important graphic, if you have one you can choose). 
And make sure to hide your code with  `echo = FALSE` unless the point you are trying to make is about the code itself.
Your results with proper output and graphics go in your presentation, this space is for a brief summary of your project.

```{r load-data, echo = FALSE}
# load data here, if you like
```

## Presentation

Our presentation can be found [here](presentation/presentation.html).

## Data 

Include a citation for your data here. 
See http://libraryguides.vu.edu.au/c.php?g=386501&p=4347840 for guidance on proper citation for datasets. 
If you got your data off the web, make sure to note the retrieval date.

## References

List any references here. You should, at a minimum, list your data source.


# Joining Datasets

```{r load-package}
library(tidyverse)
library(tibble)
library(readxl)
```

```{r}
Queer_Books_Database_1_ <- read_excel("data/Queer Books Database.xlsx", 
    sheet = "Fiction Database")
Queer_Books_Database <- read_excel("data/Queer Books Database.xlsx", 
    sheet = "Non-Fiction Database")
Queer_Books_Name_ <- read_excel("data/Queer_Books_Name .xlsx")
goodreads_top100_from1980to2023_v2_1_ <- read_csv("data/goodreads_top100_from1980to2023_v2 (1).csv")
book1_100k <- read_csv("data/book1-100k.csv")
```

```{r load-dataset}
fiction <- data_frame(Queer_Books_Database_1_)
```

```{r}
Non_fiction_book <- data_frame(Queer_Books_Database)
```

# Joining Datasets

```{r Name-Goodread}
Queer_Goodread <- Queer_Books_Name_ %>% 
  left_join(goodreads_top100_from1980to2023_v2_1_, by = c("Title" = "Book"))
```

```{r Name-Fiction}
Queer_Fiction <- Queer_Books_Name_ %>% 
  left_join(fiction, by = "Title")
```

```{r Name-Nonfiction}
Queer_nonfiction <- Queer_Books_Name_ %>% 
  left_join(Non_fiction_book, by = "Title")
```

```{r Join-all}
Queer_all <- Queer_Goodread %>% 
  inner_join(Queer_Fiction, by = "Title") %>% 
  inner_join(Queer_nonfiction, by = "Title")
```

```{r}
Queer_Book1 <- Queer_Books_Name_ %>% 
  left_join(book1_100k, by = c("Title" = "Name"))
```

```{r}

```



