---
title: "Project proposal"
author: "Queer Lit"
output: github_document
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
library(readr)
library(stringr)
library(readxl)
library(dplyr)
library(hexbin)
library(visdat)
library(naniar)
library(UpSetR)
```

## 1. Introduction

General Research Question: Examining the correlation between the number of Reviews a book receives and the Rating. Understanding a possible evolution of queer literature from the 1970s - 2024.

Examining the general relation between reviews and rating from the Goodreads rating system. The data has been formulated from Goodreads lists and can be found here:

Queer_books-data - https://docs.google.com/spreadsheets/d/11cubO21f_MJb85e4IrpsaNH9Sm4-6B8-kL2CnUNGGdE/edit  
Kandi Grey Queer Lit Shelf - https://www.goodreads.com/review/list/59261401-kandi-grey?shelf=queer-books_all 

The variables include id (Id number of book from), Title, Author (Last name, First name), Additional Authors, ISBN, ISBN13, Average Rating ( 1- 5), Num Rating (Number of Ratings), Publisher, Binding (Paperback, Hardcover, eBook, ect), Number of Pages, Year Published, Original Publication Year, Exclusive Shelf (File name). 
500 observations have been recorded. 
Goodreads is a site and a tool used by people to categorise books they have written and read. Goodreads has different ways to help users categorise books; I created a ‘bookshelf’ to collect books from Lists that other Goodreads users supplied. I collected the top 500 from several Queer Literature lists to create the dataset we are using. Using this data, we can examine the correlation between the Average Rating and the Number of Reviews received, if the Binding method is related to the year published, if there is a common publisher for Queer literature. 
Possible Further Study: Examining the relation to genre within Queer Literature. The differences across the queer spectrum.

Vocabulary Definitions: 
Queer: does not fall under the heterosexual/romantic spectrum, i.e. falls under the LGBTQIA+ spectrum
Queer Literature Definition: literature either written by an author that identifies as queer, the themes of the book itself are queer related, either through character, theme, topic ect. 


## 2. Data

```{r load-data}

queer_books_data <- read_excel("../data/queer-books_data.xlsx")
book_genre <- read_excel("../data/data_genre (1).xlsx")
```

Codebook

id: Book number

title: Book title

author: Book author

author l-f: Book author last name and then first name

additional_authors: Additional author beside the first author

ISBN: Book identifier number

ISBN_13: Book identifier number with 13-digits long

average_rating: Book average rating 

num_rating: Total number of rating 

publisher: Book publisher

binding: Book binding type

number_of_pages: Number of pages of the book

year_published: Book’s year published (first and/or latest publication)

original_publication_year: First publication year

exclusive-shelf:file name

```{r glimpse-data-frame}
glimpse(queer_books_data)
```


## 3. Ethics review
Limitations: 
The data is subjective, as it comes from Kandi's personal Goodreads profile and contains personal bias. It also a selection of 500 books from different queer book lists provided on Goodreads, which contains human bias for literature. 
The Goodreads system has a voting and review system that relies on human interaction, and therefore human subjectivity. 
The lack of information generated to fully comprehend the data is also limited in that there are variables not included within the data collected e.g. Country of origin, languages published in, gender of author - which would be significant to include when discussing the point of representation within literature. 
The are also limitations within the form of data collected: literature. The variables do not include a labeling system for Adult, New Adult, Young Adult, or Children. The genres given are broad sense of what the books contains and not specifics e.g. the label Fantasy includes Sci-fi and High Fantasy which don't share many common literary aspects other than both being Fantasy. 

Positive Effects: 
There is a lack of presented queer data regarding literature and this data would be a stepping point for future projects to understand how representation is important within communities - the data covers a spectrum of queer representation but it only a sample. Some of the queer identities include: asexuality, aromantic, gay, lesbian, trans and more. The positive effect of this is the initiative to start, to foster curiosity and awareness of these subjects. 
If this project to were be expanded using full Goodreads lists (or attempting to categorical analyse every book ever uploaded) rather than a sample from Kandi's personal profile, the impact would be useful in understanding the trends within literature over time, if there has been an increase, speculation on the relationship between queer authors and the works that they write, and the links to a more international community. 

Negative Effects: 
As the data is taken from a personal account, the negative repercussions (if there are any) would fall directly on Kandi due to personal connection. 
There is a possibility for misinterpretations within the real world due to the limitations placed on the data mentioned above. 

Minimizing Negative Impacts: 
Through communicating clearly what the data includes and does not include, how the personal bias affects the data collected, hopefully the points of misinterpretations and misconstructions can be minimized. 
This project is supposed to represent the possibilities of analysis that can be done within the literary world from a queer perspective, and the specifics within queer literature. 



## 4. Data analysis plan

```{r corelation-rating-numrating}
queer_books_data %>% 
  ggplot(aes( x = num_ratings, y = average_rating)) +
  geom_hex() +
  scale_fill_viridis_b() +
  labs( title = "Correlation of average rating and number of rating",
        subtitle = "Goodread data from 1974 - 2024",
        x = "Number of rating",
        y = "Average rating")
```

```{r published-dacade}
queer_books_data <- queer_books_data %>% 
  mutate(year_by_decade = case_when(year_published %in% c("1970","1971", "1972", "1973", "1974","1975", "1976", "1977", "1978", "1979") ~ "70s",
                                    year_published %in% c("1980","1981", "1982","1983","1984","1985","1986","1987","1988","1989") ~ "80s",
                                    year_published %in% c("1990","1991","1992","1993","1994","1995","1996","1997","1998","1999") ~ "90s",
                                    year_published %in% c("2000","2001","2002","2003","2004","2005","2006","2007","2008","2009") ~ "2000s",
                                    year_published %in% c("2010","2011","2012","2013","2014","2015","2016","2017","2018","2019") ~ "2010s",
                                    year_published %in% c("2020","2021","2022","2023","2024") ~ "2020s"))
```

```{r original-publishing-dacade}
queer_books_data <- queer_books_data %>% 
  mutate(originalyear_by_decade = case_when(original_publication_year %in% c("1897","1890", "1872","-380","-1200") ~ "<1900s",
                                            original_publication_year %in% c("1910","1911", "1912","1913","1914","1915","1916","1917","1918","1919") ~ "10s",
                                            original_publication_year %in% c("1920","1921", "1922","1923","1924","1925","1926","1927","1928","1929") ~ "20s",
                                            original_publication_year %in% c("1930","1931", "1932","1933","1934","1935","1936","1937","1938","1939") ~ "30s",
                                            original_publication_year %in% c("1940","1941", "1942","1943","1944","1945","1946","1947","1948","1949") ~ "40s",
                                            original_publication_year %in% c("1950","1951", "1952","1953","1954","1955","1956","1957","1958","1959") ~ "50s",
                                            original_publication_year %in% c("1960","1961", "1962","1963","1964","1965","1966","1967","1968","1969") ~ "60s",
                                            original_publication_year %in% c("1970","1971", "1972", "1973", "1974","1975", "1976", "1977", "1978", "1979") ~ "70s",
                                    original_publication_year %in% c("1980","1981", "1982","1983","1984","1985","1986","1987","1988","1989") ~ "80s",
                                    original_publication_year %in% c("1990","1991","1992","1993","1994","1995","1996","1997","1998","1999") ~ "90s",
                                    original_publication_year %in% c("2000","2001","2002","2003","2004","2005","2006","2007","2008","2009") ~ "2000s",
                                    original_publication_year %in% c("2010","2011","2012","2013","2014","2015","2016","2017","2018","2019") ~ "2010s",
                                    original_publication_year %in% c("2020","2021","2022","2023","2024") ~ "2020s"))
```


```{r decade-published-graph}
queer_books_data %>% 
  ggplot(aes( x = year_by_decade)) +
  geom_bar()
```

```{r first-published-gragh}
queer_books_data %>% 
  ggplot(aes( x = originalyear_by_decade)) +
  geom_bar()
```


```{r}
queer_books_data %>%
  ggplot(aes( x = year_published, fill = year_published)) +
  geom_bar() +
  scale_fill_viridis_d()
```

```{r republished-book}
republished_book <- queer_books_data %>% 
  filter(year_published != original_publication_year)

#ggplot(data = republished_book,
 #      mapping = aes( x = year_published, y = original_published)) +
  #geom_point() ##what would be a good geom to visualize this data. 
```

```{r}
queer_books_data %>%
  ggplot(aes(x = fct_rev(fct_infreq(binding)), fill = binding)) +
  geom_bar() +
  scale_fill_viridis_d() +
  guides(fill = "none") +
  coord_flip()+
  labs(title = "Binding types's frequency",
       x = NULL, y = NULL) 
```

Look at which publisher has more number of book published. 

```{r publisher-sum-books}
queer_books_data %>% 
  count(publisher) %>% 
  arrange(desc(n))
```

Look at the books' title published by the top five of most frequent publishers. Then look at the correlation of these books' average ratings and number of ratting.

```{r books-of-publisher-sum}
Books_and_publishers <- queer_books_data %>% 
  filter(publisher %in% c("Less Than Three Press", "Vintage", "Dreamspinner Press", "Little, Brown Books for Young Readers", "CreateSpace Independent Publishing Platform", "DAW", "Spectra", "Tor Books")) 

ggplot(data = Books_and_publishers,
       mapping = aes( y = publisher, fill = "average_rating")) +
  geom_bar() +
  scale_fill_viridis_d() +
  labs(title = "Publishers and their books rating",
       subtitle = "Correlation of books in top publishers and their average rating",
       x = NULL, y = NULL)
```

```{r}
book_genre <- book_genre %>% 
  select("title", "genre_1", "genre_2", "genre_3", "genre_4", "genre_5")


queer_books_genre <- queer_books_data %>% 
  left_join(book_genre, by = c("title" = "title"))

queer_books_genre %>% 
  ggplot(aes(x = fct_rev(fct_infreq(genre_1)), fill = genre_1)) +
  geom_bar() +
  coord_flip() +
  scale_fill_viridis_d() +
  guides(fill = "none") +
  labs( x = "genre")
```


```{r missing-data}
vis_miss(queer_books_genre)
```




