---
title: "Project proposal"
author: "Team name"
output: github_document
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
library(readr)
library(stringr)
library(readxl)
library(dplyr)
```

## 1. Introduction

General Research Question: 

## 2. Data

```{r load-data}

queer_books_data <- read_excel("../data/queer-books_all.xlsx")

```

Codebook

id: Book number

Title: Book title

Author: Book author

Author l-f: Book author last name and then first name

Additional authors: Additional author beside the first author

ISBN: Book identifier number

ISBN 13: Book identifier number with 13-digits long

Average Rating: Book average rating 

Num Rating: Total number of rating 

Publisher: Book publisher

Binding: Book binding type

Number of Pages: Number of pages of the book

Year Published: Book’s year published (first and/or latest publication)

Original Publication Year: First publication year

Exclusive sheft: 

```{r glimpse-data-frame}
glimpse(queer_books_data)
```


## 3. Ethics review

## 4. Data analysis plan

```{r}
queer_books_data %>%
  ggplot(aes( x = `Year Published`)) +
  geom_bar()
  
```

```{r}
queer_books_data <- queer_books_data %>% 
  mutate(republished_book = !is.na(`Year Published` %in% `Original Publication Year`))
  
```

